---
title: "2503_aRcgis"
output: html_document
---

# ArcGIS in R  

ArcGIS is a set of powerful propriety geospatial tools developed by ESRI. 
It is widely used in the industry and academia. 
In this blog, I will explore the `arcgis` R package by getting data from
the Australian tree crop map database, hosted on ArcGIS.

### Austlralian tree crop map

The [Australian tree crop map](https://www.une.edu.au/research/research-centres-institutes/applied-agricultural-remote-sensing-centre/collaborative-r-and-d-opportunities/industry-applications-and-maps/australian-tree-crops) 
is a database of tree crops in Australia.
The map is the work of the [Applied Agricultural Remote Sensing Centre](https://www.une.edu.au/research/research-centres-institutes/applied-agricultural-remote-sensing-centre) 
at the University of New England.  

The data is hosted on ArcGIS and can be accessed at the following link:
https://www.arcgis.com/home/item.html?id=12bdb57eec7548dd90409603645cfd5e  


### Installation  

Full documentation can be found [at the following link on installation and setup.](https://developers.arcgis.com/r-bridge/installation/)
This is hosted on R-universe.

`arcgis` is a meta-package for `arcgisutils`, `arcgislayers`, `arcgisgeocode` 
and `arcgisplaces`  

```{r eval = FALSE}
install.packages("arcgis", repos = c("https://r-arcgis.r-universe.dev", "https://cloud.r-project.org"))
```

Next lets load the library into the R environment.  

```{r}
library(arcgis)
library(sf) # also needed for working with spatial data
```

### Query features  

The URL to query point features from the Australian Tree crop is as follows:
https://services5.arcgis.com/3foZbDxfCo9kcPwP/ArcGIS/rest/services/AARSC_treecrop_point/FeatureServer/0  

```{r}
atcm_pts <- arc_open("https://services5.arcgis.com/3foZbDxfCo9kcPwP/ArcGIS/rest/services/AARSC_treecrop_point/FeatureServer/0")
atcm_pts
```

Here we see this feature server contains point geometry, with a coordinate reference
system 7844 and has Query capabilities.  

We can obtain more metadata information from the feature server using the `arc_meta` function.

```{r}
list_fields(atcm_pts)
```


### Selecting data  

The package contains This returns a `sf` simple features object.  

We can query the fields in the `sf` object using the `list_fields` function.  

The where argument provides a feature to filter the data based on SQL syntax.

```{r}
arc_select(atcm_pts, where = c("commodity = 'avocado'"))
```


#### Query data by spatial reference

From a 10km radius around a point.
I will choose the centre of Alstonville in Northern New South Wales.  

```{r}
# Define a point (e.g., coordinates for a location)
point_coords <- c(153.4404,-28.8421)  # Example coordinates (longitude,latitude)

# Create a spatial points data frame
# We will use the crs 7844 specified in the ATCM metadata
alston <- st_sfc(st_point(c(point_coords)), crs = 7844)

# Buffer the point to create a polygon (e.g., 10 kilometers)
buffer_alston <- st_buffer(alston, 10000)

# transform the data to WGS84 so we can plot it on a leaflet map
buffer_alston <- st_transform(buffer_alston, 4326)


library(leaflet)

leaflet(buffer_alston) |>
  addProviderTiles(provider = "Esri.WorldImagery")|>
  addMarkers(lng = point_coords[1], lat = point_coords[2], popup = "Alstonville")|>
  addPolygons(color = "green", weight = 2, fillOpacity = 0.2)
```


Now lets query and download only the data from inside the buffer polygon.

```{r}
alston_farms <- 
  arc_select(atcm_pts, 
             filter_geom = buffer_alston,
             predicate = "intersects")

# transform the data to WGS84 so we can plot it on a leaflet map
alston_farms <- st_transform(alston_farms, 4326)

leaflet(alston_farms) |>
  addProviderTiles(provider = "Esri.WorldImagery")|>
  addCircleMarkers(color = "orange", 
                   weight = 1, 
                   fillOpacity = 0.8,
                   radius = 4,
                   popup = alston_farms$commodity)|>
  addMarkers(lng = point_coords[1], lat = point_coords[2], 
             popup = "Alstonville")
  
```

Lets do this again, however instead of point data, lets get the polygon data from
the following link:
https://services5.arcgis.com/3foZbDxfCo9kcPwP/ArcGIS/rest/services/AARSC_treecrop_polygon/FeatureServer/0


```{r}
atcm_poly <- arc_open("https://services5.arcgis.com/3foZbDxfCo9kcPwP/ArcGIS/rest/services/AARSC_treecrop_polygon/FeatureServer/0")
list_fields(atcm_poly)
```


Now lets query and download only the data from inside the buffer polygon.

```{r}
# lets get the data from inside a bounding box of the buffer polygon
buffer_box <- 
  st_as_sfc(st_bbox(buffer_alston))

# request only the data in the box
alston_farms <- 
  arc_select(atcm_poly, 
             filter_geom = buffer_box,
             predicate = "intersects")

# transform the data to WGS84 so we can plot it on a leaflet map
alston_farms <- st_transform(alston_farms, 4326)

leaflet(alston_farms) |>
  addProviderTiles(provider = "Esri.WorldImagery")|>
  addPolygons(color = "orange", 
                   weight = 1, 
                   fillOpacity = 0.8,
                   popup = alston_farms$commodity)|>
  addPolygons(data = buffer_box, color = "green", weight = 2, fillOpacity = 0.2)
  
```








